---
title: "respo script"
format: html
---
####### Respo Code for P:R Rates on Seawater ######
#### Adpted from Respo Code for PI Light and Dark Runs #####
### Created by: Nyssa Silbiger; Last updated on: 2025-07-26 ###
## Adpated 2025-04-26 ###

############## Introduction to code/script ####################
## this script will help us process the raw data gathered during respirometry runs. 
## need to change for specific project/experimental variables 

### Install Packages #####
## if these packages are not yet installed, install them 
## great for updates or new users 
```{r}
if ("segmented" %in% rownames(installed.packages()) == 'FALSE') install.packages('segmented')
if ("plotrix" %in% rownames(installed.packages()) == 'FALSE') install.packages('plotrix')
if ("gridExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('gridExtra')
if ("LoLinR" %in% rownames(installed.packages()) == 'FALSE') devtools::install_github('colin-olito/LoLinR')
if ("chron" %in% rownames(installed.packages()) == 'FALSE') install.packages('chron')
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse')
if ("here" %in% rownames(installed.packages()) == 'FALSE') install.packages('here')
if ("patchwork" %in% rownames(installed.packages()) == 'FALSE') install.packages('patchwork')
if ("PNWColors" %in% rownames(installed.packages()) == 'FALSE') install.packages('PNWColors')

```

#Read in required libraries
##### Include Versions of libraries
```{r}
library(segmented)
library(plotrix)
library(gridExtra)
library(LoLinR)
library(lubridate)
library(chron)
library(patchwork)
library(tidyverse)
library(here)
library(PNWColors)
library(ggrepel)
library(reshape2)
library(viridis)
library(car)
library(future)
library(furrr)
library(dplyr)
```


############# now it's time to code ############
################################################
# get the file path
#set the path to all of the raw oxygen datasheets
## these are saved onto the computer in whatever file path/naming scheme you saved things to 
#you can change to individual run folders if needed
```{r}
path.p<-here("data","respofiles","RawO2") 
```
#the location of all your respirometry files

# bring in all of the individual files
```{r}
filenames_final<-basename(list.files(path = path.p, pattern = "csv$", recursive = TRUE)) 
```
#list all csv file names in the folder and subfolders

#basename above removes the subdirectory name from the file, re-name as file.names.full
```{r}
file.names.full<-list.files(path = path.p, pattern = "csv$", recursive = TRUE) 
```

#empty chamber volume
```{r}
ch.vol <- 500 #mL #of small chambers 
```

######### Load and tidy files ###############
############################################
#Load your respiration data file, with all the times, water volumes(mL), #not doing dry weight just SA
#RespoMeta <- read_csv(here("Data","RespoFiles","Respo_Metadata_SGDDilutions_Cabral_Varari.csv"))
```{r}
BioData <- read_csv(here("data","respofiles","Fragment_Measurements_Vest.csv"))
RespoMeta <- read_csv(here("data","respofiles","Respo_meta.csv"))
#View(BioData)
```

# join the data together

```{r}
Sample_Info <- left_join(RespoMeta, BioData)
#View(Sample_Info)
```

##### Make sure times are consistent ####
# make start and stop times real times, so that we can join the respo output and sample_info data frames
```{r}
Sample_Info <- Sample_Info %>% 
  #drop_na(sample_ID) %>% 
  unite(date,start_time,col="start_time",remove=F, sep=" ") %>% 
  unite(date,stop_time,col="stop_time",remove=F, sep=" ") %>%
  mutate(start_time = mdy_hms(start_time)) %>% 
  mutate(stop_time = mdy_hms(stop_time)) %>% 
  mutate(date = mdy(date))
```



#generate a 4 column dataframe with specific column names
# data is in umol.L.sec
```{r}
n_light_levels<-2 # number of unique light levels

RespoR <- tibble(.rows =length(filenames_final)*n_light_levels,
                 sample_ID = NA,
                 Intercept = NA,
                 umol.L.sec = NA,
                 Temp.C = NA,
                 Light_level = NA,
                 Light_value = NA,
                 run_block = NA)
```

######### Create a for loop! ###############
############################################
```{r}
for(i in 1:length(filenames_final)) {
  FRow <- as.numeric(which(Sample_Info$FileID_csv==filenames_final[i])) # stringsplit this renames our file
 
   Respo.Data1 <- read.csv(skip=1,file.path(path.p, paste0(file.names.full[i]))) %>% # reads in each file in list
    dplyr::select(Date, Time, Value, Temp) %>% # keep only what we need: Time stamp per 1sec, Raw O2 value per 1sec, in situ temp per 1sec
    unite(Date,Time,col="Time",remove=T, sep = " ") %>%
    drop_na() %>% 
    mutate(Time = mdy_hms(Time)) #%>% # convert time
    
   #mutate(help = i) ##if stuck in forloop with error from filter, can check RespoR and see at what row the forloop stopped working  

  ## cut the data by start and stop times from metadata
  #Use start time of each light step from the metadata to separate data by light stop
  
  oxy_subsets <- Sample_Info[FRow,] %>%
      pmap(function(Light_level, start_time, stop_time, ...) {
        data <- Respo.Data1  %>%
          filter(Time >= start_time & Time <= stop_time) %>%
          arrange(Time) %>%
          mutate(t_sec = as.numeric(difftime(Time, first(Time), units = "secs"))) %>% #keep everything in seconds
          mutate(Light_level = Light_level) %>%
          filter(t_sec > 120) %>%                          # drop first 2 min (120 s)
          filter(row_number() %% 10 == 0)                  # keep every 10th row
      #   data <- Respo.Data1  %>%
      #   filter(Time >= start_time & Time <= stop_time) %>%
      #   mutate(sec = row_number()) %>%# add an id for each row to help remove the first few mins
      #   mutate(Light_level = Light_level,
      #          sec = sec) %>%
      #   filter(sec > 60)  %>%# delete the first 2 mins of data assuming freq of 0.5 Hz
      #   mutate(row_number = row_number()) %>%
      #   filter(row_number %% 10 == 0) %>%  # keep every 10th row only to thin the data
      #   select(-row_number) %>%
      #   mutate(sec2 = row_number())  #update the row numbers
      # #return(subset)
    }) 
   
 
  # Combine into one long dataframe with ID labels
  combined_oxy <- bind_rows(oxy_subsets)
     
  # Get the filename without the .csv
  rename<- sub("_O2.csv","", filenames_final[i])
  
    ### plot and export the thinned data ####
  p1<- ggplot(combined_oxy, aes(x = t_sec, y = Value)) +
    geom_point(color = "dodgerblue") +
    labs(
      x = 'Time (seconds)',
      y = expression(paste(' O'[2],' (',mu,'mol/L)')),
      title = "original"
    )+
    facet_wrap(~Light_level, scales = "free_y")
  
  
  ##Olito et al. 2017: It is running a bootstrapping technique and calculating the rate based on density
  #option to add multiple outputs method= c("z", "e "pc")
  
  # Define function for fitting LoLinR regressions to be applied to all intervals for all samples
  fit_reg <- function(data) {
    rankLocReg(xall = data$t_sec, yall = data$Value, 
               alpha = 0.2, method = "pc", verbose = FALSE)
  }
  
   # Setup for parallel processing
  future::plan(multisession)
  
  # Map LoLinR function onto all intervals of each sample's thinned dataset
  df <- combined_oxy %>%
    dplyr::select(t_sec, Value, Light_level, Temp)%>%
    mutate(t_sec = as.numeric(t_sec))%>%
    nest_by(Light_level) %>%
    ungroup()%>%
    mutate(regs = furrr::future_map(data, fit_reg), # run the LOLinR fit in parallel
           Temp.C = map_dbl(map(data, "Temp"), mean),# get the mean temperature
           RegStats =map(regs, function(x){ # extract the intercept and slope for the parameters
             x$allRegs %>%
               slice(1) %>%
               select(Intercept = b0,
                      umol.L.sec = b1)
           }) )

    
 #  Plot regression diagnostics
 
  for(j in 1:length(df$Light_level)){
   pdf(paste0(here("output"),"/",rename,"_",j,".pdf" ))
    plot(df$regs[[j]])
    dev.off()
  }
  # pdf(paste0(here("Output"),"/",rename,".pdf"))
  # plot(regs)
  # dev.off() 
  
  df<-df %>%
    dplyr::select(Light_level,Temp.C, RegStats ) %>%
    unnest(RegStats) %>%
    mutate(sample_ID = rename) %>%
    left_join(Sample_Info[FRow,] %>%
                select(sample_ID, Light_level, Light_value, run_block)) # make sure the light value (or whatever other metadata you want) is in the final DF
    
  ################################
  # fill in all the O2 consumption and rate data
  
  RespoR[FRow,"Temp.C"]<-df$Temp.C
  RespoR[FRow,"sample_ID"]<-df$sample_ID
  RespoR[FRow,"Intercept"]<-df$Intercept
  RespoR[FRow,"umol.L.sec"]<-df$umol.L.sec
  RespoR[FRow,"Light_level"]<-df$Light_level
  RespoR[FRow,"Light_value"]<-df$Light_value
  RespoR[FRow,"run_block"]<-df$run_block

 }  
```


```{r}
 #export raw data and read back in as a failsafe 
#this allows me to not have to run the for loop again !!!!!
write_csv(RespoR, here("data","respofiles","Respo_RUN2.csv"))  
```

```{r}
RespoR <- read_csv(here("data","respofiles","Respo_RUN2.csv"))
```

