---
title: "respo script"
format: html
---
####### Respo Code for P:R Rates on Seawater ######
#### Adpted from Respo Code for PI Light and Dark Runs #####
### Created by: Nyssa Silbiger; Last updated on: 2025-07-26 ###
## Adpated 2025-04-26 ###

############## Introduction to code/script ####################
## this script will help us process the raw data gathered during respirometry runs. 
## need to change for specific project/experimental variables 

### Install Packages #####
## if these packages are not yet installed, install them 
## great for updates or new users 
```{r}
if ("segmented" %in% rownames(installed.packages()) == 'FALSE') install.packages('segmented')
if ("plotrix" %in% rownames(installed.packages()) == 'FALSE') install.packages('plotrix')
if ("gridExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('gridExtra')
if ("LoLinR" %in% rownames(installed.packages()) == 'FALSE') devtools::install_github('colin-olito/LoLinR')
if ("chron" %in% rownames(installed.packages()) == 'FALSE') install.packages('chron')
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse')
if ("here" %in% rownames(installed.packages()) == 'FALSE') install.packages('here')
if ("patchwork" %in% rownames(installed.packages()) == 'FALSE') install.packages('patchwork')
if ("PNWColors" %in% rownames(installed.packages()) == 'FALSE') install.packages('PNWColors')

```

#Read in required libraries
##### Include Versions of libraries
```{r}
library(segmented)
library(plotrix)
library(gridExtra)
library(LoLinR)
library(lubridate)
library(chron)
library(patchwork)
library(tidyverse)
library(here)
library(PNWColors)
library(ggrepel)
library(reshape2)
library(viridis)
library(car)
library(future)
library(furrr)
library(dplyr)
```


############# now it's time to code ############
################################################
# get the file path
#set the path to all of the raw oxygen datasheets
## these are saved onto the computer in whatever file path/naming scheme you saved things to 
#you can change to individual run folders if needed
```{r}
path.p<-here("data","respofiles","RawO2", "RUN2") 
```
#the location of all your respirometry files

# bring in all of the individual files
```{r}
filenames_final<-basename(list.files(path = path.p, pattern = "csv$", recursive = TRUE)) 
```
#list all csv file names in the folder and subfolders

#basename above removes the subdirectory name from the file, re-name as file.names.full
```{r}
file.names.full<-list.files(path = path.p, pattern = "csv$", recursive = TRUE) 
```

#empty chamber volume
```{r}
ch.vol <- 500 #mL #of small chambers 
```

######### Load and tidy files ###############
############################################
#Load your respiration data file, with all the times, water volumes(mL), #not doing dry weight just SA
#RespoMeta <- read_csv(here("Data","RespoFiles","Respo_Metadata_SGDDilutions_Cabral_Varari.csv"))
```{r}
BioData <- read_csv(here("data","respofiles","Fragment_Measurements_Vest.csv"))
RespoMeta <- read_csv(here("data","respofiles","Respo_meta.csv"))
#View(BioData)
```

# join the data together

```{r}
Sample_Info <- left_join(RespoMeta, BioData)
#View(Sample_Info)
```

##### Make sure times are consistent ####
# make start and stop times real times, so that we can join the respo output and sample_info data frames
```{r}
Sample_Info <- Sample_Info %>% 
  #drop_na(sample_ID) %>% 
  unite(date,start_time,col="start_time",remove=F, sep=" ") %>% 
  unite(date,stop_time,col="stop_time",remove=F, sep=" ") %>%
  mutate(start_time = mdy_hms(start_time)) %>% 
  mutate(stop_time = mdy_hms(stop_time)) %>% 
  mutate(date = mdy(date))
```



#generate a 4 column dataframe with specific column names
# data is in umol.L.sec
```{r}
n_light_levels<-2 # number of unique light levels

RespoR <- tibble(.rows =length(filenames_final)*n_light_levels,
                 sample_ID = NA,
                 Intercept = NA,
                 umol.L.sec = NA,
                 Temp.C = NA,
                 Light_level = NA,
                 Light_value = NA,
                 run_block = NA)
```

######### Create a for loop! ###############
############################################
```{r}
###forloop#####
for(i in 1:length(filenames_final)) {
  FRow <- as.numeric(which(Sample_Info$FileID_csv==filenames_final[i])) # stringsplit this renames our file
 
   Respo.Data1 <- read.csv(skip=1,file.path(path.p, paste0(file.names.full[i]))) %>% # reads in each file in list
    dplyr::select(Date, Time, Value, Temp) %>% # keep only what we need: Time stamp per 1sec, Raw O2 value per 1sec, in situ temp per 1sec
    unite(Date,Time,col="Time",remove=T, sep = " ") %>%
    drop_na() %>% 
    mutate(Time = mdy_hms(Time)) #%>% # convert time
    #mutate(help = i) ##if stuck in forloop with error from filter, can check RespoR and see at what row the forloop stopped working  
```

## cut the data by start and stop times from metadata
#Use start time of each light step from the metadata to separate data by light stop

```{r}
 oxy_subsets <- Sample_Info[FRow,] %>%
      pmap(function(Light_level, start_time, stop_time, ...) {
        data <- Respo.Data1  %>%
          filter(Time >= start_time & Time <= stop_time) %>%
          arrange(Time) %>%
          mutate(t_sec = as.numeric(difftime(Time, first(Time), units = "secs"))) %>% #keep everything in seconds
          mutate(Light_level = Light_level) %>%
          filter(t_sec > 120) %>%                          # drop first 2 min (120 s)
          filter(row_number() %% 10 == 0)                  # keep every 10th row
      #   data <- Respo.Data1  %>%
      #   filter(Time >= start_time & Time <= stop_time) %>%
      #   mutate(sec = row_number()) %>%# add an id for each row to help remove the first few mins
      #   mutate(Light_level = Light_level,
      #          sec = sec) %>%
      #   filter(sec > 60)  %>%# delete the first 2 mins of data assuming freq of 0.5 Hz
      #   mutate(row_number = row_number()) %>%
      #   filter(row_number %% 10 == 0) %>%  # keep every 10th row only to thin the data
      #   select(-row_number) %>%
      #   mutate(sec2 = row_number())  #update the row numbers
      # #return(subset)
    }) 
```

